---
title: "Pear Assessment 1 for reproducible Research"
author: "S. Brugere"
date: "10 mai 2017"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
  
## Loading and preprocessing the data
```{r read data}
if (! file.exists("activity.csv")) unzip("activity.zip")
data <- read.csv("activity.csv", header=T, sep=",")
```
  
  
## What is mean total number of steps taken per day?
```{r calc total steps by day, fig.width=12}  
par(mfrow=c(1,1))
totalByDay <- aggregate(steps~date,data=data,FUN="sum")
with(totalByDay, hist(steps,breaks=12))
abline(v = mean(totalByDay$steps, na.rm=T), col = "blue", lwd=2)
abline(v = median(totalByDay$steps), col ="red", lwd=2, lty=2)
```  
  
The mean total number of steps taken per day is `r sprintf("%5.0f",mean(totalByDay$steps, na.rm=T))`.  
The median total number of steps taken per day is `r median(totalByDay$steps)`.  
  
## What is the average daily activity pattern?
```{r means per interval, fig.width=12, fig.height=5}
means <- aggregate(steps~interval, data = data, FUN = "mean", na.rm = TRUE)
plot(steps~interval, data = means, type="l", main="Average number of steps by interval",ylim=c(0,300))
abline(v = means[which.max(means$steps),1], col = "red", lwd = 2)
```
  
The interval which corresponds to the max average number of steps is `r means[which.max(means$steps),1]`.  
    
## Imputing missing values
```{r count missing values}
cnt <- sum(is.na(data$steps))
```  
The number of missing values is `r cnt`.  

```{r imputing missing values, fig.width=12}
data2 <- transform(data, steps = ifelse(is.na(steps), means[,2], data[,1]))
totalByDay <- aggregate(steps~date, data=data2,FUN="sum")
with(totalByDay, hist(steps,12))
abline(v = mean(totalByDay$steps, na.rm=T), col = "blue", lwd=2)
abline(v = median(totalByDay$steps), col ="red", lwd=2, lty=2)
```  
The strategy with missing values is to replace NA values with the average of the interval.
The only difference with the first histogram is the frequency of the interval of the mean value?  
  
## Are there differences in activity patterns between weekdays and weekends?
```{r type of day, fig.width=12,fig.height=10}
data3 <- transform(data, typeOfDay = 
    ifelse(weekdays(as.Date(data$date,"%Y-%m-%d")) %in% c("samedi","dimanche"), "weekend", "weekday"))
means <- aggregate(steps~interval+typeOfDay, data = data3, FUN = "mean", na.rm = TRUE)
par(mfrow=c(2,1))
plot(steps~interval, data = means[means$typeOfDay=="weekday",], type = "l", 
     main="Average number of steps by interval\n Week Days", cex.main = 1.5, ylim=c(0,300))
plot(steps~interval, data = means[means$typeOfDay=="weekend",], type = "l", 
     main="Average number of steps by interval\n Week End", cex.main= 1.5, ylim=c(0,300))

```  
